---
title: "Creating a dataset"
output: html_notebook
---
## Data structures
R has a wide variety of objects for holding data, including _scalars_, _vectors_, _matrices_, _arrays_, _data frames_, and _lists_.

### Vectors

- Vectors are one-dimensional arrays that can hold numeric data, character data, or logical data.

- The combine function _c()_ is used to form vector.

- The data in a vector must be of only one type or mode (numeric, character or logical). 

__Numeric vector__
```{r}
a <- c(1, 2, 5, 3, 6, -2, 4)
a
```
__Character vector__
```{r}
b <- c("one", "two", "three")
b
```

__Logical vector__
```{r}
c <- c(TRUE, TRUE, TRUE, FALSE, TRUE, FALSE)
c
```
Unlike programming languages such as C++, Java, and Python, R position indices start with 1 rather than 0.

```{r}
a <- c("k", "j", "h", "a", "c", "m")
```

```{r}
a[3]
```
```{r}
a[c(1, 3, 5)]
```
```{r}
a[2:6]
```
The colon operator __:__ used in the last statement generates a sequence of numbers. For example, _a <- c(2:6)_ is equivalent to _a <-(2, 3, 4, 5, 6)_.
