---
title: "Creating a dataset"
output: html_notebook
---
## Data structures
R has a wide variety of objects for holding data, including _scalars_, _vectors_, _matrices_, _arrays_, _data frames_, and _lists_.

### Vectors

- Vectors are one-dimensional arrays that can hold numeric data, character data, or logical data.

- The combine function _c()_ is used to form vector.

- The data in a vector must be of only one type or mode (numeric, character or logical). 

__Numeric vector__
```{r}
a <- c(1, 2, 5, 3, 6, -2, 4)
a
```
__Character vector__
```{r}
b <- c("one", "two", "three")
b
```

__Logical vector__
```{r}
c <- c(TRUE, TRUE, TRUE, FALSE, TRUE, FALSE)
c
```
Unlike programming languages such as C++, Java, and Python, R position indices start with 1 rather than 0.

```{r}
a <- c("k", "j", "h", "a", "c", "m")
```

```{r}
a[3]
```
```{r}
a[c(1, 3, 5)]
```
```{r}
a[2:6]
```
The colon operator __:__ used in the last statement generates a sequence of numbers. For example, _a <- c(2:6)_ is equivalent to _a <-(2, 3, 4, 5, 6)_.


### Matrices

- A _matrix_ is a two-dimensional array in which each element has the same mode (_numeric_, _character_, or _logical_).

- Matrices are created with the _matrix_ function.

- The general format is: 
```
mymatrix <- matrix(vector, nrow=number_of_rows, ncol=number_of_columns, byrow=logical_value, dimnames=list(char_vector_rownames, char_vector_colnames))
```

* _vector_: Contains the elements for the matrix.
* _nrow_ and _ncol_: Specify the row and columns dimensions.
* _dimnames_: Contains optional row and column labels stored in character vectors.
* _byrow_: Indicates whether the matrix should be filled by row (_byrow=TRUE_) or by column (byrow=FALSE). The default is by column.

__Creating matrices__

_Creates a 5 X 4 matrix_
```{r}
y <- matrix(1:20, nrow=5, ncol=4)
y
```


```{r}
cells <- c(1, 26, 24, 68)
rnames <- c("R1" , "R2")
cnames <- c("C1", "c2")
```

_2 X 2 matrix filled by rows_
```{r}
mymatrix <- matrix(cells, nrow=2, ncol=2, byrow=TRUE, dimnames=list(rnames, cnames))
mymatrix
```
__Using matrix subscripts__
```{r}
x <- matrix(1:10, nrow=2)
x
```
Select the elements in the second row.
```{r}
x[2,]
```
Select the elements in the second column.
```{r}
x[,2]
```
Select the element in the first row and fourth column
```{r}
x[1,4]
```
Select the elements in the first row and the fourth and fifth columns.
```{r}
x[1,c(4, 5)]
```

